### 2016-01-22 writeups/msr_workshop/paper.tex
Section 2 mods

\documentclass[a4paper,twocolumn]{article}

\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}

\newcommand{\ignore}[1]{}

\title{
  It's Never Too Late To Learn\\
  \emph{Using Online Machine Learning in Distributed Systems}
}
\author{}
\date{}

\begin{document}
\maketitle

\section{Motivation}

Building a modern distributed service that guarantees service-level objectives
is hard. Complexity is introduced by the deployment environment and exacerbated
as the service evolves with time to accommodate hardware/environment changes or
support new features.
For example, to build a simple caching layer that guarantees a certain hit rate,
response latency, and throughput, developers need to take into account the
expected workload accessing the cache, the number of machines available to run
it, the network connecting these machines, the storage hierarchy within each
machine, and interference from other applications running on the network and
machines.  A change in any one of these factors could render the service
woefully inefficient.

Traditionally, this complexity is addressed by using highly-skilled developers
to build custom systems that are optimized for specific configurations and
objective functions.  Is there a general, principled, approach for dealing with
complexity in distributed systems? Hand-tuning the design and configuration of a
system does not scale well with increased complexity or changes in software
design, hardware architecture, deployment environment, and service-level
objectives.

We propose to use online learning to cope with this complexity. Specifically, we
view a {\em decision} made by a distributed system in the framework of
contextual online learning with {\em partial feedback} (where the learner only
gets to observe the result of the action it took, rather than all possible
actions). This is a natural fit for distributed systems where: there is plenty
of context surrounding each decision, decisions are made online, and feedback is
received only for decisions that were executed.  Thus, a hand-designed policy in
a distributed systems can be replaced with policy that has been optimized by a
contextual learning system. By supplying rich contexts to this policy,
developers capture the complexity of our environment without having to
understand it. Additionally, by optimizing the policy continuously, the learning
system can swiftly adapt to changes in the environment or service.

%% Hussam: I changed "full information" to "full feedback" to stay consistent
%% with the naming above. Confirmed with Akshay that this is correct terminology

Machine learning has certainly been used in the past to optimize system
decisions. Most of this use has been restricted to either offline settings or
expect {\em full feedback}, whereby a supervised learner is trained on data
that has been annotated with correct labels.
Some systems work in an online setting with full feedback. An example of that
is a caching layer that uses a learning system to dynamically switch between
existing (hand-designed) cache eviction policies, by running all policies in
parallel and measuring their performance.
Systems that work in an online setting with partial feedback
tend to be for news or ad recommendation, but examples from systems
infrastructure exist. For example, caching dynamic query content [], tuning
parameters of multicore data structures [], allocating resources to cluster
computing tasks [], and allocating servers to web applications [].  In all of
these systems, the context used to make a decision, and the decision itself, are
{\em local} to a machine. This is achieved by either devising a decentralized
algorithm in the first place, or giving up on global optimality. We are not
aware of any system that makes ``global'' decisions based on truly distributed
architecture or state.

%% Hussam: I don't know what is this "and yet" referencing. So I am taking it out.
%%
%%And yet we use hand-designed policies to make these decisions all the time, such
%%as for request routing, replica placement, replica selection, cloud resource
%%allocation, failure recovery tasks, and others.

What is it about distributed systems that makes it difficult to apply online
learning at decision points such as request routing, replica placement, cloud
resource allocation, and failure recovery? This proposal identifies fundamental
challenges posed by distributed systems that must be analyzed and addressed in
order to make online learning a standard tool for distributed infrastructure
systems.  These challenges hold even if we assume that the online learning
system is a black-box that delivers optimal decisions with zero latency. In
other words, they are fundamental to the way we design distributed systems.  We
begin by highlighting some of the factors that contribute to these challenges,
and then list the challenges themselves.

\subsection*{Known unknowns}

Systems often make decisions without complete knowledge of the factors that
affect, or could affect, those decisions.  For example, a caching system evicts
items based on what it believes future requests will look like, without actually
knowing the future.  In other cases, incomplete knowledge is due to the absence
of data-computation locality.  For example, a load balancer makes decisions
based on what it thinks is the current load on other machines, e.g. based on
data it collected recently, rather than the actual load at that very instant in
time.

In both of these situations, assumptions are made about the unknowns offline
during \emph{design time}, resulting in different request routing and caching
algorithms. A change to the \emph{run-time} environment could easily render
these choices suboptimal.

\ignore{
Online learning has the potential to make these decisions online, based on current contexts,
and can be nimble in adapting to any changes to the run-time
environment.
}

\subsection*{Unknown unknowns}

In some situations, system designers have no way of knowing deployment details
that might impact a system's performance. For example, a system could be built
for external deployment and used by third-party customers, or it could be
deployed in a multi-tenant environment where co-located services could affect
its performance in unexpected ways. In these cases, it is nearly impossible for
developers to account for all unknown variables at design time.

\ignore{
However, an ML agent is able to make
decisions online based on inputs from the run-time environment.
}
\ignore{
To be clear, ML has been used in some distributed systems. However, it has not
been used in a general way to make online decisions about core distributed
systems implementation details. We advocate that developers should focus on
high-level objectives of their system, and use online ML to make to compute the
specifics (such as policies for request routing, replica placement, or failure
monitoring) at run-time based on observed metrics.
}

\section{Using Online Learning in Distributed Systems}

As mentioned earlier, we adopt the framework of contextual online learning with
partial feedback.  To separate the systems concerns we are trying to highlight
from machine learning concerns, we abstract the learning system as a black-box
and assume it provides optimal decisions with zero latency. For a given decision
over a discrete action space, the interface to the learning system is as
follows: given a {\em context} that summarizes the state of the world, the
learning system chooses an {\em action} to take, which our system does and later
reports a {\em reward} indicating how good the action was. An example of a
learning system that supports this interface is the Decision
Service~\footnote{http://aka.ms/mwt}. Now consider a load balancer using that
learning system to make its request-routing decisions. The context for each
decision could be load information of each machine in the cluster, the action
could be a specific machine to route the request to, and the reward could be the
latency of processing the request.

Thus we have reduced our concerns to generating an appropriate context,
executing an action (told to us by the learning system), and reporting reward
information.  We believe that fundamental properties of distributed systems make
even these tasks difficult, as described below.

\subsection*{Locality}

Decisions made by an ML agent in distributed system are likely to rely on input
data collected from diverse and physically distributed components. As a result,
the input to the learning loop might be stale or inconsistent. What could be
done to mitigate those issues while making reasonable real-time decisions?
Additionally, the decisions outputted by the learning agent need to be executed
by diverse and distributed components. How could routine distributed systems
issues such as message delays in an asynchronous setting affect the learning
system?

\subsection*{Hierarchy}
Distributed systems are often designed in a hierarchical manner, where
subcomponents are monitoring and managing others. These components and layers of
indirection simplify reasoning about the system at large. However, it is not
clear how hierarchy will affect an ML agent.
\begin{itemize}
  \item Will hierarchy aid the learning system via parallelism? Or will it hinder
    it since not all features are visible at every level?
  \item Should learning happen in each component of the system? Or should it be
    centralized?
  \item If learning is happening at multiple levels of the hierarchy, will that
    create interference between the learners? Or will they complement each
    other?
   \item Should feature vectors cut across the levels or not?
\end{itemize}

%% Sid:
%% In general should be composable if at each level can define decision and reward,
%% and exploration randomization is different at each level

\subsection*{Failures}

Failures are an expected occurrence in every distributed system. How do learning
systems cope with data omission or inability to execute decisions due to
failures? Can the ML agent learn from incomplete or potentially corrupt data?
How sensitive should the learner be to transient conditions in the system?

%% What features of the run-time system are important to the ML agent? How can they
%% be captured effectively? Can the ML agent learn from incomplete data? For
%% example, if an ML agent needs to make decisions regarding performance
%% optimization, would ignoring garbage collection activities or the VMM result in
%% corrupting the learned model? How sensitive should the learner be?
%% 
%% - quesiton of what features to include is there, but being conservative usually the right
%% way to go
%% - bigger quesiton is how ml component deals with omissions and other problems. Separate
%% into failure sections? Yeah, replace this with ``failures'' and move feature selection stuff
%% down to last section

%% Hussam: Sid, I don't understand what you mean by randomization in this context
%% So I am taking this subsection out.
%% \subsection*{Randomization}
%% 
%% An ML agent can be trained offline with pre-labeled data. However, is this
%% possible in the context of distributed systems? An online learning systems is
%% more logical in this context, however, how do we quantify the costs of
%% exploration vs exploitation of learned strategies?
%% 
%% - Probably omit this, or replace with ``Randomization''. Basically, we already know
%% that exploratoin-exploitation is the right approach, and that will be encapsulated
%% by the ML component (its decisions will reflect this, and it will take care of
%% computing probabilities etc.). But there's an interesting point about inherent randomization
%% in systems
%%

\section{Expanding the scope}

There are two drawbacksTwo drawbacks: opacity and scope.

Opacity (also, which features matter?)
- We are also pursuing work on cracking models open to see what they are doing inside.
There are different ways to gain insight as to what a model is doing. For example a simple
linear model may learn a vector of weights on the supplied features. These weights can be
studied, and indeed some techniques such as lasso are oftne used to reduce the number of features. However, these
techniques need to be handled carefully. Just becasue a feature has 0 weight doesn't mean it
can be removed: perhaps it is unimportant preceisely because it exists, and its removal all of
asudden changes the weight on other features.

\ignore{
contextual multi-armed bandits. idea is to provide a context, get an action,
and only observe reward for that action. This matches the dist systems world well. It will generate new
policies using ML that we could not come up with by just thinking, and may not
even be able to describe. We argue in favor of opacity and rely instead on the guarantees
provided by the ML optimization.
}

Scope:
The above replaces the policy behind a decision with one backed by an ML learning system.
Indeed systems like hte DS were designed to intervene at the decision making point in a
slim way. The above outlines the issues which tackled would result in ML based policies,
new ones. However, once the above has been mastered, there is the question of whether ML
can influence distributed systems design at a higher level. For example, can it suggest
different mechanisms for achieving a goal, or perhaps a different design of the system in the first place?

\end{document}