### 2016-01-24 writeups/msr_workshop/paper.tex
completed and cleaned up

\documentclass[letterpaper,twocolumn]{article}
%\documentclass[pdftex,twocolumn,10pt,letterpaper]{article}
\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}

\newcommand{\ignore}[1]{}

\title{
  It's Never Too Late to Learn:\\
  \emph{Using Online Machine Learning in Distributed Systems}
}
\author{Hussam Abu-Libdeh and Siddhartha Sen \\ \small{Microsoft Research NYC}}
\date{}

\begin{document}
\maketitle

\section{Motivation}

Building a modern distributed service that guarantees service-level objectives
is hard. Complexity is introduced by the deployment environment and exacerbated
as the service evolves with time to accommodate hardware/environment changes or
support new features. Some of the unknown variables are known: a caching layer
must cope with changing workload distributions; a load balancer might have stale
load information from the servers. But some of the unknowns are unknown: a
system could be deployed in a multi-tenant environment where co-located
services affect its performance in unexpected ways.

\ignore{
For example, to build a simple caching layer that guarantees a certain hit rate,
response latency, and throughput, developers need to take into account the
expected workload accessing the cache, the number of machines available to run
it, the network connecting these machines, the storage hierarchy within each
machine, and interference from other applications running on the network and
machines.  A change in any one of these factors could render the service
woefully inefficient.
}

Traditionally, this complexity is addressed by using highly-skilled developers
to build custom systems that are optimized for specific configurations and
objective functions.  Is there a general, principled, approach for dealing with
complexity in distributed systems? Hand-tuning the design and configuration of a
system does not scale well with increased complexity or changes in the
environment or service-level objectives.

We propose to use online learning to cope with this complexity. Specifically, we
cast the decisions made by a distributed system in the framework of
{\em contextual online learning with partial feedback}.  This is a natural fit
for many decisions where: there is plenty of context relevant to the decision, the
decision is made online, and feedback is only received for the action that
was taken.  Thus, we can replace a hand-designed policy for the decision with a
policy that has been optimized by a contextual learning system. By supplying
rich contexts to the learner, developers can capture the complexity of the
environment without having to understand it.  By training the learner
continuously, the system can adapt swiftly to changes in the environment
or service.

%% Hussam: I changed "full information" to "full feedback" to stay consistent
%% with the naming above. Confirmed with Akshay that this is correct terminology

Online learning has been used in systems before. Some of these systems require
full feedback, which is not practical, but others operate in the partial
feedback setting.  The latter systems tend to be recommendation systems (e.g.,
for personalized news or ads), but some infrastructure systems exist, such as
for caching dynamic query content or allocating resources to
cluster jobs.  In all of these systems, the contextual decisions are made {\em
locally} within a machine, either because the system is not distributed, or it
uses a decentralized algorithm, or it does not aim for global optimality.
\ignore{We are not aware of any system that makes ``global'' decisions based on truly
distributed architecture or state.}

\ignore{
Machine learning has certainly been used in the past to optimize system
decisions. Most of this use has been restricted to either offline settings or
expect {\em full feedback}, whereby a supervised learner is trained on data
that has been annotated with correct labels.
Some systems work in an online setting with full feedback. An example of that
is a caching layer that uses a learning system to dynamically switch between
existing (hand-designed) cache eviction policies, by running all policies in
parallel and measuring their performance.
Systems that work in an online setting with partial feedback
tend to be for news or ad recommendation, but examples from systems
infrastructure exist. For example, caching dynamic query content [], tuning
parameters of multicore data structures [], allocating resources to cluster
computing tasks [], and allocating servers to web applications [].  In all of
these systems, the context used to make a decision, and the decision itself, are
{\em local} to a machine. This is achieved by either devising a decentralized
algorithm in the first place, or giving up on global optimality. We are not
aware of any system that makes ``global'' decisions based on truly distributed
architecture or state.
}

%% Hussam: I don't know what is this "and yet" referencing. So I am taking it out.
%%
%%And yet we use hand-designed policies to make these decisions all the time, such
%%as for request routing, replica placement, replica selection, cloud resource
%%allocation, failure recovery tasks, and others.

More importantly, prior work does not abstract, analyze, or address the
fundamental challenges imposed by distributed systems when applying online
learning.  The purpose of this proposal is to identify these challenges and
motivate us to start working on them.  We believe these are systems
challenges: that is, they hold even if we assume black-box access to a
learning system that provides optimal decisions in real-time. By overcoming
these challenges, we stand to gain principled mechanisms and tools for
automating intelligence in our distributed infrastructure systems, such as our
cloud.

\ignore{
and points such as request routing, replica placement,
cloud resource allocation, and failure recovery? This proposal identifies fundamental
challenges posed by distributed systems that must be analyzed and addressed in
order to make online learning a standard tool for distributed infrastructure
systems.  These challenges hold even if we assume that the online learning
system is a black-box that delivers optimal decisions with zero latency. In
other words, they are fundamental to the way we design distributed systems.  We
begin by highlighting some of the factors that contribute to these challenges,
and then list the challenges themselves.
}

\ignore{
\subsection*{Known unknowns}

Systems often make decisions without complete knowledge of the factors that
affect, or could affect, those decisions.  For example, a caching system evicts
items based on what it believes future requests will look like, without actually
knowing the future.  In other cases, incomplete knowledge is due to the absence
of data-computation locality.  For example, a load balancer makes decisions
based on what it thinks is the current load on other machines, e.g. based on
data it collected recently, rather than the actual load at that very instant in
time.

In both of these situations, assumptions are made about the unknowns offline
during \emph{design time}, resulting in different request routing and caching
algorithms. A change to the \emph{run-time} environment could easily render
these choices suboptimal.

\ignore{
Online learning has the potential to make these decisions online, based on current contexts,
and can be nimble in adapting to any changes to the run-time
environment.
}

\subsection*{Unknown unknowns}

In some situations, system designers have no way of knowing deployment details
that might impact a system's performance. For example, a system could be built
for external deployment and used by third-party customers, or it could be
deployed in a multi-tenant environment where co-located services could affect
its performance in unexpected ways. In these cases, it is nearly impossible for
developers to account for all unknown variables at design time.

\ignore{
However, an ML agent is able to make
decisions online based on inputs from the run-time environment.
}
\ignore{
To be clear, ML has been used in some distributed systems. However, it has not
been used in a general way to make online decisions about core distributed
systems implementation details. We advocate that developers should focus on
high-level objectives of their system, and use online ML to make to compute the
specifics (such as policies for request routing, replica placement, or failure
monitoring) at run-time based on observed metrics.
}
}

\section{Challenges}

\ignore{We consider classic decisions made by distributed systems where the
choice of action is discrete, such as: where to route a request, how many replicas to
place (and where), or when to raise an alarm during failures.}

For a given decision, we can view the world as having two components: the
distributed system and the learning system, which we treat as a black box.  The
system interacts with the learner as follows: it provides a {\em context}
to the learner summarizing the state of the world, the learner returns 
an {\em action} to take, the system executes the action and reports a {\em
reward} to the learner indicating how good it was.  An example of a
contextual learning system that supports this interface is the Decision
Service\footnote{http://aka.ms/mwt}.  An example of a decision is which
server a load balancer should send a request to: the context could be the load
information of every server, the action is the chosen server, and
the reward could be the latency of the request.

\ignore{
We separate the world into a As mentioned earlier, we
adopt the framework of contextual online learning with partial feedback.  To separate the systems concerns we are trying to highlight
from machine learning concerns, we abstract the learning system as a black-box
and assume it provides optimal decisions with zero latency. For a given decision
over a discrete action space, the interface to the learning system is as
follows: given a {\em context} that summarizes the state of the world, the
learning system chooses an {\em action} to take, which our system does and later
reports a {\em reward} indicating how good the action was. An example of a
learning system that supports this interface is the Decision
Service~\footnote{http://aka.ms/mwt}. Now consider a load balancer using that
learning system to make its request-routing decisions. The context for each
decision could be load information of each machine in the cluster, the action
could be a specific machine to route the request to, and the reward could be the
latency of processing the request.
}

Thus the system is responsible for generating an appropriate context, executing
an action, and reporting the reward.  The following properties of distributed
systems make these tasks challenging (there may be others).

\paragraph{Locality.}
Decisions in a distributed system may rely on information that resides on
diverse and physically disparate components. This information may not be readily
available at the point of decision (e.g., the load balancer).
As a result, the context supplied to the learner may be stale or
incomplete.  How does staleness affect the learner's ability to optimize
decisions? Can we exploit ``smoothness'' in the context to mitigate this effect?
Can we eliminate staleness using hardware or software techniques?

Another disparity may occur between the point of decision and the
point of execution (e.g., the server chosen to handle a request).  This
separation prevents the decider from enforcing the action; indeed, the action
may be changed or ignored.  Contextual learning systems can tolerate
downstream logic that alters the decision, provided rewards are
correctly reported and certain independence assumptions are not
violated.  Are there distributed systems that violate these assumptions?

It is common for reward information to be reported by a separate
component. We do not expect this to pose a problem beyond the fact that delays 
in reward reporting cause delays in learning.

\ignore{typically reported separately by a asynchronously While reward
information is also not available at the decision location, this is the norm in contextual learning
settings, where reward information is only discernible after a delay and often
by a separate subsystem. Thus, we do not expect this to pose a problem. (The faster reward information is collected, the
faster the resulting data point can be used by the learning subsystem.)
}

\paragraph{Failures.}
The absence of locality results in a lack of fate sharing
between the components involved in a decision.  Failures in any of
these components affect the inputs to the learning system.
For example, server failures in the load balancing example could
result in variable staleness or even missing/corrupt information in
the context.  How should information that is known to be missing 
be conveyed to the learner?  Can corrupt data derail the learner's optimization?

Failures can also prevent a component from executing the action, causing the
system to change the action or ignore it.  Can such failures violate the
independence assumptions mentioned above?  In general, how sensitive should
the learner be to transient conditions in the system?

\paragraph{Hierarchy.}
Components in a distributed system often interact with and manage other
components. This hierarchy simplifies building, maintaining, and reasoning
about the system at large. It is not clear how to take hierarchy into
account when optimizing decisions via contextual learning.  Contextual learning
is generally composable, in that decisions of lower-level
components can safely influence those of higher-level components, and those of
upstream components can influence those of downstream components.   Should the
context for each decision be limited to its level in the hierarchy, or should it span
multiple (or all) levels?  Should the decision points roughly mimic the
hierarchy of the system, or should they cut across? Consider the example of a
deep memory hierarchy. One could imagine capturing contextual
information across the entire hierarchy and using it to decide which level to cache an item in, instead of deciding at each level in isolation and relying on evictions to move between levels.

\paragraph*{Randomization.}

Contextual learning balances exploration (a type of controlled randomization)
with exploitation so that it can discover potentially better policies
without compromising current performance.  Exploration can be completely
encapsulated by the learning system and manifested through the actions it
recommends.  However, many distributed systems already use
randomization, such as for cache eviction, load balancing, 
and replica placement.  Can this randomization be leveraged by the learning
system to achieve its exploration goals at no additional cost?

\ignore{First, if multiple components are making decisions independently, it is
important to ensure that the randomization used by each (e.g. the seed passed to
the learning system) are independent of each other.}

\ignore{
An ML agent can be trained offline with pre-labeled data. However, is this
possible in the context of distributed systems? An online learning systems is
more logical in this context, however, how do we quantify the costs of
exploration vs exploitation of learned strategies?
}

\section{Drawbacks}

The approach we have outlined has at least two drawbacks that merit their own
investigation.  The first is that we are using machine-learned policies that are
generally opaque, in that we cannot easily inspect them to learn things like
which parts of the context really matter.  Cracking open such policies is
certainly an interesting research direction, but our current view is to accept
opacity as the price we pay for offloading complexity to the learner.

The second drawback is that these policies replace hand-designed policies only
at existing decision points.  In other words, machine learning is used as an
alternative within an existing design and architecture, as opposed to 
influencing the design/architecture of the system itself.  It is very
conceivable that in the future distributed systems will be co-designed with
learners.

%% Scope:
%% The above replaces the policy behind a decision with one backed by an ML learning system.
%% Indeed systems like hte DS were designed to intervene at the decision making point in a
%% slim way. The above outlines the issues which tackled would result in ML based policies,
%% new ones. However, once the above has been mastered, there is the question of whether ML
%% can influence distributed systems design at a higher level. For example, can it suggest
%% different mechanisms for achieving a goal, or perhaps a different design of the system in the first place?

\end{document}